### 2015-05-12 paper/paper.tex
Added basic paper template

\documentclass[a4paper,11pt]{article}

\usepackage[paper=a4paper,left=30mm,width=150mm,top=25mm,bottom=25mm]{geometry}
\usepackage[hidelinks]{hyperref}    % All references will auto hyperlink
\usepackage[font={small}]{caption}  % Make figure captions smaller
\usepackage{graphicx}               % For the displaying of images in floats
\usepackage{mathtools}              % For more advanced math usage
\usepackage{poltakmacros}           % Personal macro package (https://gist.github.com/poltak/6ed59e8f30c9c399143f)

\begin{document}

\include{titlepage}                 % Pull in title page

\newpage
\pagenumbering{roman}
\tableofcontents
\newpage

\pagenumbering{arabic}

\section{Introduction} % (fold)
\label{sec:introduction}

% section introduction (end)

\newpage                            % Make references section

\bibliographystyle{acm}
\bibliography{biblio.tex}

\end{document}

### 2015-05-12 paper/paper.tex
Wrote introduction and skeleton of paper

@@
 
+While relational database management systems (RDBMS) have been somewhat of a ``go-to'' solution for for a number of years
+for general data storage and management problems that many application developers face, we have noted a recent rise
+in the use of non-relational data management tools~\cite{padhy2011rdbms}. Most of these tools have traditionally fallen into the domain of
+big data analytics, with platforms such as the Hadoop ecosystem~\footnote{https://hadoop.apache.org/} being notably popular.
+However, outside of the domain of big data, looking more at general purpose data storage and management, what are now
+commonly referred to as ``NoSQL'' solutions are proving to be a popular solution.
+
+NoSQL databases refer to those databases that are not built on top of the relational algebraic concepts, as laid out by
+Codd in 1970~\cite{codd1970relational}, unlike the more commonly used RDBMS technologies, such as
+MySQL~\footnote{https://www.mysql.com/}. Being free of the strictness the relational model enforces on its data allows
+NoSQL databases to focus less on the overall structure of data, and more on factors such as scalability and
+performance~\cite{leavitt2010will}.
+
+While the relational model is a good fit for many data problems, its strictness in terms of flexibility of managing data
+eventually led to the introduction of the NoSQL model. The following characteristics can be given as a starting point
+for NoSQL databases in comparison to relational databases~\cite{indrawan2012database}:
+
+\begin{itemize}
+  \item \textbf{Unstructured data support:} While the relational model would often force data to be stored in tabular
+  formats, the NoSQL model does not force any kind of data schema.
+  \item \textbf{Designed with distributed processing and horizontal scalability in mind:} Given the commoditisation of
+  computer hardware in the last decade, support for horizontal scaling and processing among clusters is an important
+  factor for adoption.
+  \item \textbf{Less strict adherence to ACID principles:} While the relational model attempted to very much adhere to
+  the transactional principles of data atomicty, consistency, isolation, and durability (ACID), this very much impacts
+  performance in terms of distributed computing. Relaxing the strictness of adherence to these principles, allow many
+  NoSQL databases to make the trade-off for higher performance.
+\end{itemize}
+
+Of course, these differences between the NoSQL model and relational model vary between each individual database
+technology's design, and trade-offs are often made depending on the goals and aims for that given database.
+
+In this paper, we will look at the use of MongoDB~\footnote{https://www.mongodb.org/}, a popular NoSQL database solution,
+as a solution for a case study based upon a railway data problem using data from Monash University's Institute of Railway
+Technology (IRT). An overview of the case study in question will be given in~\sectref{sec:case_study_overview}. A small
+overview of the research already in this area will be given in~\sectref{sec:research_context}. Implementation and
+evaluation details will be given in~\sectref{sec:implementation} and~\sectref{sec:evaluation_and_discussion}, respectively,
+before concluding in~\sectref{sec:conclusion}.
+
 % section introduction (end)
+
+\newpage
+
+\section{Case Study Overview} % (fold)
+\label{sec:case_study_overview}
+
+% section case_study_overview (end)
+
+\newpage
+
+\section{Research Context} % (fold)
+\label{sec:research_context}
+
+% section research_context (end)
+
+\newpage
+
+\section{Implementation} % (fold)
+\label{sec:implementation}
+
+% section implementation (end)
+
+\newpage
+
+\section{Evaluation and Discussion} % (fold)
+\label{sec:evaluation_and_discussion}
+
+% section evaluation_and_discussion (end)
+
+\newpage
+
+\section{Conclusion} % (fold)
+\label{sec:conclusion}
+
+% section conclusion (end)
 

### 2015-05-13 paper/paper.tex
Added case study overview content

@@
   \item \textbf{Less strict adherence to ACID principles:} While the relational model attempted to very much adhere to
-  the transactional principles of data atomicty, consistency, isolation, and durability (ACID), this very much impacts
+  the transactional principles of data atomicity, consistency, isolation, and durability (ACID), this very much impacts
   performance in terms of distributed computing. Relaxing the strictness of adherence to these principles, allow many
@@
 Technology (IRT). An overview of the case study in question will be given in~\sectref{sec:case_study_overview}. A small
-overview of the research already in this area will be given in~\sectref{sec:research_context}. Implementation and
+overview of the MongoDB database will be given in~\sectref{sec:mongo_overview}. Implementation and
 evaluation details will be given in~\sectref{sec:implementation} and~\sectref{sec:evaluation_and_discussion}, respectively,
@@
 
+The case study that is being looked at involves a project that has been worked at at the Monash University Institute of
+Railway technology (IRT). The project involves trains the operate in the Pilbara region of Western Australia, taking
+ore and minerals from loading points at mines to a specified unloading port. The data that is being dealt with comes from
+numerous categories of sensors being placed on certain specialised train cars to record data monitoring track and train
+car conditions~\cite{darby2003development,darby2005track}.
+
+Data currently gets unloaded at sent back in large batches to remote servers once the car completes a trip and pulls back
+into port~\cite{thomas2012taking}. The project currently makes use of a RDBMS solution to manage data storage, however
+this solution is faced with many problems that currently require painful work-arounds. The most obvious of which involves
+unreliable data being received from sensors. For example, in the case of a damaged or failed sensor, reliable data cannot
+be guaranteed to be returned from such a sensor. Hence, data received by the remote server is often inconsistently
+structured, and thus the data has to go through a series of preprocessing work arounds to fit in within the strict
+schema that the RDBMS expects. As the only guaranteed consistency in the data that gets received is the timestamp and
+geocoordinates, the strictly structured relational model is not an appropriate solution.
+
+The Monash IRT team are currently investigating further solutions in the NoSQL big data space, where they intend to
+replace the current system with an appropriately tested solution. While the scope of that project is much larger than
+what will be covered in this paper, we will look at the differences in what is possible when attempting to use a NoSQL
+database for the data storage and management.
+
+For this paper, we will be looking at the possibility of using MongoDB, a popular NoSQL database that does away with
+the concept of the schemas and tables so commonly found in databases following the relational model. This is done with
+the expectation of better handling of the given inconsistent data.
+
 % section case_study_overview (end)
@@
 
-\section{Research Context} % (fold)
-\label{sec:research_context}
+\section{MongoDB Overview} % (fold)
+\label{sec:mongo_overview}
 
-% section research_context (end)
+% section mongo_overview (end)
 
@@
 \bibliographystyle{acm}
-\bibliography{biblio.tex}
+\bibliography{biblio}
 

### 2015-05-13 queries.js
Added sample mongo queries

@@ -1,0 +1,95 @@
+// Command to insert CSV with headers to Mongo:
+//mongoimport -d railway -c train --type csv --file ~/Downloads/test.csv --headerline
+
+/*
+Runs a basic map reduce to get the total distance travelled on each track
+Note that it excludes where km value is given as less than 0.
+ */
+function totalDistanceMR() {
+  var mapFunc = function() {
+    emit(this.Track, this.km);
+  };
+
+  var reduceFunc = function(keyTrack, kms) {
+    return Array.sum(kms);
+  };
+
+  var finalizeFunc = function(key, reducedVal) {
+    reducedVal.avg = reducedVal.qty / reducedVal.count;
+
+    return reducedVal;
+  };
+
+  db.train.mapReduce(
+    mapFunc,
+    reduceFunc,
+    {
+      out: 'total_distance',
+      query: { km:
+                { $gt: 0 }
+             },
+      finalize: finalizeFunc
+    }
+  );
+}
+
+
+/*
+Iteratively restructure default documents into related fields
+ */
+function latLon() {
+  var updateDocsFunc = function(doc) {
+    // Add location nested doc
+    doc.Loc = {
+      Lat: doc.Lat,
+      Lon: doc.Lon
+    };
+
+    // Add Bounce nested doc
+    doc.Bounce = {
+      Front: doc.BounceFront,
+      Rear: doc.BounceRear
+    };
+
+    // Add Rock nested doc
+    doc.Rock = {
+      Front: doc.RockFront,
+      Rear: doc.RockRear
+    };
+
+    // Add SND nested doc
+    doc.SND = {
+      SND1: doc.SND1,
+      SND2: doc.SND2,
+      SND3: doc.SND3,
+      SND4: doc.SND4
+    };
+
+    // Add Acc nested doc
+    doc.Acc = {
+      Left: doc.AccLeft,
+      Right: doc.AccRight
+    };
+
+    // Delete old values
+    delete doc.Lat;
+    delete doc.Lon;
+    delete doc.BounceFront;
+    delete doc.BounceRear;
+    delete doc.RockFront;
+    delete doc.RockRear;
+    delete doc.SND1;
+    delete doc.SND2;
+    delete doc.SND3;
+    delete doc.SND4;
+    delete doc.AccLeft;
+    delete doc.AccRight;
+
+    // Write changes
+    db.train.save(doc);
+  };
+
+  db.train.find().snapshot().forEach(updateDocsFunc);
+}
+
+

### 2015-05-16 paper/paper.tex
Added further additions to MongoDB section

@@
 
+\textit{Note that most of the information relevant to MongoDB in this section is sourced from the official MongoDB
+manual.~\cite{The M0:online}}
+
+\subsection{Data storage} % (fold)
+\label{sub:data_storage}
+
+MongoDB can be described as a documented-oriented NoSQL database, which does not rely on schemas or tables to structure
+its data~\cite{parker2013comparing}. Data within Mongo is stored using JSON-like structured documents, called
+BSON~\cite{bsons6:online}, which allows for strongly typed data fields and binary data. BSON documents are then stored
+within ``collections'' which reside in a given ``database''~\cite{parker2013comparing}. To relate back to the relational
+model, these concepts within Mongo can be compared as shown in~\tabref{tab:mongo_to_relational}.
+
+\begin{table}[H]
+\centering
+\caption{MongoDB conceptual structures as they relate to the relational model concepts.}
+\label{tab:mongo_to_relational}
+\begin{tabular}{ | l | l | }
+
+\hline
+\rowcolor{gray!20}
+\textbf{MongoDB}  & \textbf{Relational Model} \\ \hline
+Database          & Database                  \\ \hline
+Collection        & Table                     \\ \hline
+BSON Document     & Record/Row/Tuple          \\ \hline
+BSON Key          & Column/Field              \\ \hline
+BSON Value        & Value                     \\ \hline
+
+\end{tabular}
+\end{table}
+
+While these concepts within Mongo can be related back to concepts within the relational model for ease of understanding,
+they are not exactly the same. For example, a table in the relational model is structured in such a way that all records
+within that table must adhere to. Conversely, in a Mongo collection, several differently structured BSON documents may
+coexist freely. BSON documents generally follow some sort of predefined structure for that collection, to maintain
+data structure sanity, for the user's sake, however none of this is enforced.
+
+This sort of freedom in terms of structure allows for similarly structured documents in database collections, where
+differences may only be found in that certain documents may miss keys that are present in other documents. This is
+highly rational in the way that, depending on the use-case, data may not be present for certain fields all the time.
+Thus instead of placing null or default values for missing data, Mongo allows the user to just leave them out of the
+relevant document. Note that document structure in Mongo is also dynamic in the way that it can be changed at any time;
+BSON keys and values may be added or removed from existing documents at any time.
+
+% subsection data_storage (end)
+
+\subsection{Data interaction} % (fold)
+\label{sub:data_interaction}
+
+MongoDB allows a lot of the similar type functionality in queries and data manipulation that is also expected in
+relational databases. Generally interactions with data are performed through JavaScript functions on collections, or cursors
+(subsets of documents within a collection), allowing general querying operations through use of the \texttt{find()} function,
+and standard CRUD operations through the \texttt{create()}, \texttt{update()}, and \texttt{remove()} functions.
+
+MongoDB officially provides support for many drivers allowing interaction with Mongo databases in different programming
+languages, including JavaScript, Python, and Java. As well as providing drivers, MongoDB also provides a default shell,
+called the \textit{mongo shell}, which allows general JavaScript code to be written to interact with the standard MongoDB
+API and perform ad-hoc queries. No official graphical user interface exists to interact with MongoDB, however there
+are numerous third party clients, which essentially wrap around mongo shell.
+
+MongoDB also supports more advanced data manipulation operations through use of aggregation functions. Generally, aggregation
+functions are user defined operations to perform on documents in collections to return results. These include general
+aggregate operation models, such as MapReduce, and often involve grouping documents by given keys and various queries
+to limit the amount of documents to process.
+
+% subsection data_interaction (end)
+
+\subsection{Further database features} % (fold)
+\label{sub:further_database_features}
+
+MongoDB also allows for common database operations, such as indexing, replication, and sharding. Mongo's indexing system
+supports multiple types of indexes, including single key, multi-key, geospatial, and hashed indexes. Indexes are stored
+internally using a B-Tree data structure
+
+In terms of joins, MongoDB,
+
+% subsection further_database_features (end)
+
 % section mongo_overview (end)

### 2015-05-16 queries.js
Added further queries

@@ -90,6 +90,15 @@ function latLon() {
   };
 
   db.train.find().snapshot().forEach(updateDocsFunc);
+
+  // Create a geospatial index on Loc nested document
+  db.train.createIndex({'Loc': '2d'});
+
+  // Create an index on km field
+  db.train.createIndex({'km': 1});
+
+  // Find all documents stating a value less than 0 for KM travelled (these are obviously wrong)
+  db.train.find({'km': {'$lt': 0}});
 }
 
 

### 2015-05-17 paper/paper.tex
Finished MongoDB subsection

@@
 \textit{Note that most of the information relevant to MongoDB in this section is sourced from the official MongoDB
-manual.~\cite{The M0:online}}
+manual.}~\cite{Mongo_man}
 
@@
 
-\begin{table}[H]
+\begin{table}[h]
 \centering
@@
 \hline
-\rowcolor{gray!20}
 \textbf{MongoDB}  & \textbf{Relational Model} \\ \hline
@@
 
-MongoDB also allows for common database operations, such as indexing, replication, and sharding. Mongo's indexing system
-supports multiple types of indexes, including single key, multi-key, geospatial, and hashed indexes. Indexes are stored
-internally using a B-Tree data structure
+\subsubsection{Lack of joins} % (fold)
+\label{ssub:joins}
 
-In terms of joins, MongoDB,
+MongoDB, like many other NoSQL databases, does not support the joins of multiple tables or, in MongoDB's case, documents,
+as is common in relational databases.
+Instead, MongoDB allows two main ways of relating different documents to each other. These are via referencing and
+nested documents. Nested documents are the main recommended way of join multiple documents, unless the documents have
+many-to-many relationships. To allow nested documents, a BSON value can store a whole other document, or arrays of documents.
+Example use-cases of where nested documents are appropriate include documents that have natural ``child'' documents that
+only it will ever need to reference. For example, a blog post document may have many nested comment documents as the
+comments are specific to a given blog post. For documents where many things may refer to it, it is recommended to use
+references via BSON values containing other document IDs.
+
+% subsubsection joins (end)
+
+\subsubsection{Indexing} % (fold)
+\label{ssub:indexing}
+
+Indexing in MongoDB is supported in a similar ways that are present in relational databases. Mongo builds a B-Tree data
+structure on a particular collection's key, allowing all documents within that collection with that key to be accessed
+in $O(log n)$ time, as opposed to $O(n)$ time if unindexed. MongoDB also supports numerous types of indexes, such as
+single-key, multi-key, geospatial, and hashed indexes.
+
+% subsubsection indexing (end)
+
+\subsubsection{Replication} % (fold)
+\label{ssub:replication}
+
+MongoDB allows replication and management of multiple copies of the same data for the purpose of increasing data
+redundancy and availability. Replication in Mongo is managed using a primary instance of data, with secondaries replicating
+the data in the primary using operation logs (\textit{oplogs}) to mimic any operations on data that were performed in the
+primary. Replicas have fault tolerance in the way that if the primary fails, an election is started to promote any of the
+secondaries to become a primary. In general, a client will only interact with a primary, however operations can also be
+specified to be performed on specific secondaries. However, as replication data is asynchronously managed, any reads
+performed specifically on secondaries may not give up-to-date data.
+
+% subsubsection replication (end)
+
+\subsubsection{Sharding} % (fold)
+\label{ssub:sharding}
+
+Sharding in MongoDB allows for horizontal scaling of a Mongo database, where a given database or collection is shared
+among different ``shards'', or multiple MongoDB instances, generally running on separate machines. Sharding is performed
+on an indexed document key, dividing documents containing that key based on either range or computed hashes. Ranged-based
+sharding is generally recommended where data is generally expected to be accessed in some sort of sequence, while hash-based
+sharding offers a guaranteed random distribution of data across shards, and more balanced equally-sized shards.
+
+% subsubsection sharding (end)
 

### 2015-05-18 paper/paper.tex
Added data table

@@
 
+Implementation of the database for the given Monash IRT data was all performed using MongoDB version 3.0.3, the latest
+version as of 2015-05-18. Interaction with the MongoDB database was all performed using the mongo shell and corresponding
+official NodeJS MongoDB driver.
+
+\subsection{Overview of the data} % (fold)
+\label{sub:overview_of_the_data}
+
+The test data used for this case study was acquired from the Monash IRT team, and was an extract of real data they have
+received from the railway sensors. As only one set of data was given, simulated datasets based upon the extract have
+also been created to simulate the type of inconsistencies the IRT team are facing, including differently structured
+datasets and datasets with missing data (in the case of non-functional sensors).
+
+The dataset received from the IRT team has the layout as shown in~\tabref{tab:irt_data}. It consists of 99999 data points
+recorded at different times. The data received in this dataset is all consistent and contains readings from all the sensors.
+As this is not the case in with real data over time, simulated data has been used as stated previously. The only
+guaranteed data to be received is the \textbf{Time}, \textbf{Lat}, and \textbf{Lon} sensor recordings.
+
+\begin{table}[h]
+\caption{The different types of data received from the Monash University Institute of Railway Technology team.}
+\label{tab:irt_data}
+\tabcolsep=0.11cm
+\scalebox{0.7}{
+\begin{tabular}{ | l | l | l | }
+
+\hline
+\textbf{Data name}      & \textbf{Possible values}               & \textbf{Meaning} \\ \hline
+\textbf{Time}           & Floating point value                   & Time in milliseconds since... \\ \hline
+\textbf{Speed}          & Floating point value                   & Speed in km/h (?)  \\ \hline
+\textbf{LoadEmpty}      & -1, 0                                  & Boolean flag denoting whether or not load is present in train cars.  \\ \hline
+\textbf{km}             & Floating point value                   & Amount of kilometres travelled since...  \\ \hline
+\textbf{Lat}            & Floating point value from -90 to +90   & Geographic latitude coordinate at given reading.  \\ \hline
+\textbf{Lon}            & Floating point value from -180 to +180 & Geographic longitude coordinate at given reading.  \\ \hline
+\textbf{Track}          & Integer value from 1 to 213841         & Given track in which the train is located.   \\ \hline
+\textbf{SND1}           & Floating point value                   & ?  \\ \hline
+\textbf{SND2}           & Floating point value                   & ?  \\ \hline
+\textbf{SND3}           & Floating point value                   & ?  \\ \hline
+\textbf{SND4}           & Floating point value                   & ?  \\ \hline
+\textbf{CouplerForce}   & Floating point value                   & The amount of force couple detected in the train car.  \\ \hline
+\textbf{LateralAccel}   & Floating point value                   & The amount of lateral acceleration detected in the train car.  \\ \hline
+\textbf{AccLeft}        & Floating point value                   & The amount of acceleration detected by the left sensor.  \\ \hline
+\textbf{AccRight}       & Floating point value                   & The amount of acceleration detected by the right sensor.  \\ \hline
+\textbf{BounceFront}    & Floating point value                   & The amount of bounce detected at the front of the car.  \\ \hline
+\textbf{BounceRear}     & Floating point value                   & The amount of bounce detected at the rear of the car.  \\ \hline
+\textbf{RockFront}      & Floating point value                   & The amount of rock detected at the front of the car.  \\ \hline
+\textbf{RockRear}       & Floating point value                   & The amount of rock detected at the front of the car.  \\ \hline
+
+\end{tabular}}
+\end{table}
+
+% subsection overview_of_the_data (end)
+
 % section implementation (end)

### 2015-05-19 queries.js
Added Mongo update for converting to bools

@@ -37,8 +37,11 @@ function totalDistanceMR() {
 /*
 Iteratively restructure default documents into related fields
  */
-function latLon() {
+function preProcessing() {
   var updateDocsFunc = function(doc) {
+    // Change LoadEmpty attribute to boolean type
+    doc.LoadEmpty = doc.LoadEmpty === -1 ? false : true;
+
     // Add location nested doc
     doc.Loc = {
       Lat: doc.Lat,
@@ -101,4 +104,34 @@ function latLon() {
   db.train.find({'km': {'$lt': 0}});
 }
 
+function fixBools() {
+  // Change LoadEmpty = -1 to true
+  db.train.update(
+    // query
+    { "LoadEmpty" : 0 },
 
+    // update
+    { $set: {LoadEmpty: true} },
+
+    // options
+    {
+      "multi" : true,  // update only one document
+      "upsert" : false  // insert a new document, if no existing document match the query
+    }
+  );
+
+  // Change LoadEmpty = -1 to false
+  db.train.update(
+    // query
+    { "LoadEmpty" : -1 },
+
+    // update
+    { $set: {LoadEmpty: false} },
+
+    // options
+    {
+      "multi" : true,  // update only one document
+      "upsert" : false  // insert a new document, if no existing document match the query
+    }
+  );
+}

### 2015-05-19 paper/paper.tex
Many additions to Implementation sect

@@
 
-The dataset received from the IRT team has the layout as shown in~\tabref{tab:irt_data}. It consists of 99999 data points
-recorded at different times. The data received in this dataset is all consistent and contains readings from all the sensors.
-As this is not the case in with real data over time, simulated data has been used as stated previously. The only
-guaranteed data to be received is the \textbf{Time}, \textbf{Lat}, and \textbf{Lon} sensor recordings.
+The dataset received from the IRT team has the layout as shown in~\tabref{tab:irt_data}. It consists of 99\,999 data points
+recorded at different times, and is received as raw data before being processed into a CSV format for easy manipulation
+and subsequent storage into their database.
 
@@
 \hline
-\textbf{Data name}      & \textbf{Possible values}               & \textbf{Meaning} \\ \hline
-\textbf{Time}           & Floating point value                   & Time in milliseconds since... \\ \hline
-\textbf{Speed}          & Floating point value                   & Speed in km/h (?)  \\ \hline
-\textbf{LoadEmpty}      & -1, 0                                  & Boolean flag denoting whether or not load is present in train cars.  \\ \hline
-\textbf{km}             & Floating point value                   & Amount of kilometres travelled since...  \\ \hline
+\textbf{Name}           & \textbf{Data type}                     & \textbf{Description} \\ \hline
+\textbf{Time}           & Floating point value                   & Time in seconds since the start of the test. \\ \hline
+\textbf{Speed}          & Floating point value                   & Speed in kilometres per hour. \\ \hline
+\textbf{LoadEmpty}      & -1, 0, 1                               & 1 if load is present, 0 if empty, or -1 if unknown/in-port.  \\ \hline
+\textbf{km}             & Floating point value                   & Elapsed kilometres travelled since the start of the test.\\ \hline
 \textbf{Lat}            & Floating point value from -90 to +90   & Geographic latitude coordinate at given reading.  \\ \hline
 \textbf{Lon}            & Floating point value from -180 to +180 & Geographic longitude coordinate at given reading.  \\ \hline
-\textbf{Track}          & Integer value from 1 to 213841         & Given track in which the train is located.   \\ \hline
-\textbf{SND1}           & Floating point value                   & ?  \\ \hline
-\textbf{SND2}           & Floating point value                   & ?  \\ \hline
-\textbf{SND3}           & Floating point value                   & ?  \\ \hline
-\textbf{SND4}           & Floating point value                   & ?  \\ \hline
+\textbf{Track}          & Integer value from 1 to 213\,841         & Given track in which the train is located.   \\ \hline
+\textbf{SND1}           & Floating point value                   & Spring nest deflection count in millimetres at car corner.  \\ \hline
+\textbf{SND2}           & Floating point value                   & Spring nest deflection count in millimetres at car corner.  \\ \hline
+\textbf{SND3}           & Floating point value                   & Spring nest deflection count in millimetres at car corner.  \\ \hline
+\textbf{SND4}           & Floating point value                   & Spring nest deflection count in millimetres at car corner.  \\ \hline
 \textbf{CouplerForce}   & Floating point value                   & The amount of force couple detected in the train car.  \\ \hline
@@
 
+This dataset contains data from all the possible sensors. As previously stated, the datasets received are not always
+consistent, hence not all of these fields will always be present, or may be present and contain incorrect data readings.
+The only sensor data guaranteed to be received is the \textbf{Time}, \textbf{Lat}, and \textbf{Lon} sensor recordings.
+
 % subsection overview_of_the_data (end)
+
+\subsection{Overview of data handling} % (fold)
+\label{sub:overview_of_data_handling}
+
+None of the data returned from the sensors contain overly complicated structures, and hence MongoDB's BSON data storage
+format can easily handle all of the given data types. All given types can be handled using the BSON primitive types of
+Double and 32-bit Integer for floating point and integer data, respectively.
+
+MongoDB comes with a built-in tool, called \texttt{mongoimport}, for the importing of multiple common data files, including
+CSV. Hence, the IRT data can be imported into a running MongoDB instance using the following command pattern:
+
+%TODO: \texttt{mongoimport -d DATABASE_NAME -c COLLECTION_NAME --type csv --file FILE_NAME.csv --headerline}
+
+The \texttt{-d} and \texttt{-c} flags enable the specifying of database and collection within that database in which to
+import the data into. The \texttt{--headerline} flag asks Mongo to use the CSV header to automatically generate BSON
+keys for each document. So each row in the CSV file will be converted to a MongoDB document, and stored in the specified
+database collection. By default, MongoDB tries to guess the data type for each value, with numbers represented as Doubles.
+Of course, storing every document as a flat container for different floating point values is not optimal, and pre-processing
+can be performed on Mongo to better structure the documents for later querying and processing.
+
+% subsection overview_of_data_handling (end)
+
+\subsection{Updating of data types} % (fold)
+\label{sub:restructuring_of_data}
+
+One of the first operations needed to be performed to restructure the data within MongoDB is to assign better data types
+fields where it is appropriate. With the given sensor data fields, the most obvious would be the \textbf{Track} and
+\textbf{LoadEmpty} fields, as the possible values they should be able to store are limited. With the \textbf{Track} field,
+the only values that should be present are integers representing the current track that the train car is located. Currently,
+the track numbers in the scope of the IRT project range from 1 to 213\,841. These all fall into the 32-bit integer representation
+that BSON supports, hence rather than the default Double type, it more be more appropriate to change it to the 32-bit Integer
+type. This can be done with the following code:
+
+%TODO: code
+
+With the \textbf{LoadEmpty} field, the only possible values are -1, 0, and 1, each representing different states. While
+this would be a perfect case to use an enumerated type, MongoDB does not natively support enums. The two main choices
+for handling enums is by using either the String or 32-bit Integer types. As the String type allows more expressive
+values and can be indexed to have integer-like performance, it would be most appropriate to change the field to be
+represented as a String type, mapping the existing numbers to appropriate strings. As the existing \textbf{LoadEmpty}
+value will need to be changed, this operation can be performed with the \texttt{db.collection.update()} function, as
+follows:
+
+%TODO: code
+
+Note that three different updates need to be performed on the collection, for each different value of \textbf{LoadEmpty},
+as filtered out by the query argument. As each update will need to query the collection to retreive a subset of documents
+to operate on, it would be wise to create a MongoDB index on the \textbf{LoadEmpty} key. This single-key index can be
+created using the following \texttt{db.collection.createIndex()} function:
+
+%TODO: code
+
+% subsection restructuring_of_data (end)
+
+\subsection{Restructuring of documents} % (fold)
+\label{sub:restructuring_of_documents}
+
+As the documents now contain appropriately typed key-value pairs, further modifications to the documents can be performed
+to better structure the documents. By default, the CSV importer structures the documents generated from CSV rows as flat
+documents, with all row values being represented as a discrete key-value pair. The default flat structure can be seen as
+follows:
+
+%TODO: JSON example
+
+From looking at and understanding the
+sensor data that is being dealt with, numerous sensors can be separated into distinct groups. For example, each of the
+\textbf{SND1}, \textbf{SND2}, \textbf{SND3}, and \textbf{SND4} values are from the same type of sensor located at different
+parts of the train car. Hence, rather than having four discrete key-pairs, these values can be grouped together in their
+own sub-document, or object, and pointed to by a new \textbf{SND} key. This is a more sane and appropriate way of
+structuring these values for later queries and operations.
+
+Numerous other sensor values can be grouped together in this manner for more appropriate and saner structuring using
+sub-documents within the original BSON documents. Often nested documents are used where separate tables would be used in
+a relational modelling of the dataset, which also stretches to this case, where a separate table could possible hold all
+SND sensor values.
+
+Further restructuring can be done to group the \textbf{Bounce}, \textbf{Rock}, and \textbf{Acc} values into sub-documents
+containing key-value pairs for each, finally ending up with the following document structure:
+
+%TODO: grouped JSON
+
+These restructurings could be performed with the following code, in mongo shell for example:
+
+%TODO: resturcturing code
+
+Finally, one more restructure could be performed on the existing document structure, which is grouping the \textbf{Lat}
+and \textbf{Lon} values into a single sub-document, which could be named \textbf{Loc}. This further enables future
+geospatial indexes being more easily created on the location data, for location-dependent queries and operations.
+This restructuring can be performed by the following code:
+
+%TODO: code for Loc
+
+Finally, the collection contains documents in a similar structure to:
+
+%TODO: final JSON structure
+
+% subsection restructuring_of_documents (end)
+
+\subsection{Handling missing values} % (fold)
+\label{sub:handling_missing_values}
+
+For the test datasets missing data for particular sensors, MongoDB handles these cases. The datasets include
+CSV files with different value layouts, and often missing values in random rows, apart from the time and location fields
+as they are guaranteed.
+
+For test datasets that are organised in different orders of fields, as Mongo's inner data storage format, BSON, does
+not have any ordering of its keys, documents are still generated with the appropriate data. For datasets with missing
+values in certain rows, the BSON document that is generated simply does not contain a key-value pair for the missed
+attribute. This is opposed to setting the key to a null value, or default value. Of course, a user may generate keys
+for the missing values and point them to null or default values, however there is no advantage to the extra processing.
+
+When operations are performed on multiple documents in Mongo, if the documents contain a different amount of keys, Mongo
+effectively ignores this and continues to perform the operation on all documents. Generally, specific keys and values
+are passed to operations such as these, hence if the key used does not exist within certain documents, those documents
+simply get ignored from the operation. For example, if a collection contains 100 documents, where only four of those
+documents contain the \textbf{isAdmin} key and an operation is called that depends on the key \textbf{isAdmin}, only
+those four documents within the collection will be operated on. This allows completely differently structured documents
+to be contained within the same collection (however, in a well-structured dataset, multiple collections would be used).
+
+Hence, in the IRT case study, if certain sensors break down during operation and are unable to send back data, MongoDB
+will still allow queries and operations to be performed on the overall sensor data. However, if any operations depend on
+a particular sensor's data, such as the operations performed in the pre-processing of the data
+in~\sectref{sub:restructuring_of_data}, the documents missing those sensor's data will simply be left out of the operation. This
+avoids errors that are currently encountered with the currently set default values for broken sensors. For instance, if
+a sensor is broken and needs to be used in an operation, a lot less documents will be needed to be involved in the
+query, increasing query performance. Furthermore, any reducing operations performed on the datasets, will not be subject
+to involving unneeded default values in their calculations.
+
+% subsection handling_missing_values (end)
+
+\subsection{Filtering bad data} % (fold)
+\label{sub:filtering_bad_data}
+
+% subsection filtering_bad_data (end)
 

### 2015-05-20 queries.js
Fixed and restructured broken queries

@@ -1,5 +1,5 @@
 // Command to insert CSV with headers to Mongo:
-//mongoimport -d railway -c train --type csv --file ~/Downloads/test.csv --headerline
+//mongoimport -d railway -c train --type csv --file ~/Downloads/test.csv --headerline --ignoreBlanks
 
 /*
 Runs a basic map reduce to get the total distance travelled on each track
@@ -25,23 +25,72 @@ function totalDistanceMR() {
     reduceFunc,
     {
       out: 'total_distance',
-      query: { km:
-                { $gt: 0 }
-             },
+      query: { km: { $gt: 0 }},
       finalize: finalizeFunc
     }
   );
 }
 
+function fixTypes() {
+  db.train.find({Track: {$exists: true}}).forEach(function(doc) {
+    // Cast Track to be a Int32 type
+    doc.Track = new NumberInt(doc.Track);
+
+    // Write changes to DB
+    db.train.save(doc);
+  });
+
+  // Change LoadEmpty = 1 to 'loaded'
+  db.train.update(
+    // query
+    { "LoadEmpty" : 1 },
+
+    // update
+    { $set: {LoadEmpty: 'loaded'} },
+
+    // options
+    {
+      "multi" : true,  // update only one document
+      "upsert" : false  // insert a new document, if no existing document match the query
+    }
+  );
+
+  // Change LoadEmpty = 0 to 'empty'
+  db.train.update(
+    // query
+    { "LoadEmpty" : 0 },
+
+    // update
+    { $set: {LoadEmpty: 'empty'} },
+
+    // options
+    {
+      "multi" : true,  // update only one document
+      "upsert" : false  // insert a new document, if no existing document match the query
+    }
+  );
+
+  // Change LoadEmpty = -1 to 'unknown'
+  db.train.update(
+    // query
+    { "LoadEmpty" : -1 },
+
+    // update
+    { $set: {LoadEmpty: 'unknown'} },
+
+    // options
+    {
+      "multi" : true,  // update only one document
+      "upsert" : false  // insert a new document, if no existing document match the query
+    }
+  );
+}
 
 /*
 Iteratively restructure default documents into related fields
  */
-function preProcessing() {
+function changeStructure() {
   var updateDocsFunc = function(doc) {
-    // Change LoadEmpty attribute to boolean type
-    doc.LoadEmpty = doc.LoadEmpty === -1 ? false : true;
-
     // Add location nested doc
     doc.Loc = {
[…]

### 2015-05-24 paper/paper.tex
Added code snippets

@@
 \usepackage{poltakmacros}           % Personal macro package (https://gist.github.com/poltak/6ed59e8f30c9c399143f)
+\usepackage{listings}
+\usepackage{color}
+\definecolor{lightgray}{rgb}{.9,.9,.9}
+\definecolor{darkgray}{rgb}{.4,.4,.4}
+\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
+
+\lstdefinelanguage{JavaScript}{
+  keywords={typeof, new, true, false, catch, function, delete, return, null, catch, switch, var, if, in, while, do, else, case, break},
+  keywordstyle=\color{blue}\bfseries,
+  ndkeywords={class, export, boolean, throw, implements, import, this},
+  ndkeywordstyle=\color{darkgray}\bfseries,
+  identifierstyle=\color{black},
+  sensitive=false,
+  comment=[l]{//},
+  morecomment=[s]{/*}{*/},
+  commentstyle=\color{purple}\ttfamily,
+  stringstyle=\color{red}\ttfamily,
+  morestring=[b]',
+  morestring=[b]"
+}
+
+\lstset{
+   language=JavaScript,
+   backgroundcolor=\color{lightgray},
+   extendedchars=true,
+   basicstyle=\footnotesize\ttfamily,
+   showstringspaces=false,
+   showspaces=false,
+   numbers=left,
+   numberstyle=\footnotesize,
+   numbersep=9pt,
+   tabsize=2,
+   breaklines=true,
+   showtabs=false,
+   captionpos=b
+}
 
@@
 overview of the MongoDB database will be given in~\sectref{sec:mongo_overview}. Implementation and
-evaluation details will be given in~\sectref{sec:implementation} and~\sectref{sec:evaluation_and_discussion}, respectively,
+discussion details will be given in~\sectref{sec:implementation} and~\sectref{sec:discussion}, respectively,
 before concluding in~\sectref{sec:conclusion}.
@@
 
-%TODO: \texttt{mongoimport -d DATABASE_NAME -c COLLECTION_NAME --type csv --file FILE_NAME.csv --headerline}
+\begin{lstlisting}[language=bash]
+mongoimport -d DATABASE_NAME -c COLLECTION_NAME --type csv --file FILE_NAME.csv --headerline --ignoreBlanks
+\end{lstlisting}
 
@@
 that BSON supports, hence rather than the default Double type, it more be more appropriate to change it to the 32-bit Integer
-type. This can be done with the following code:
+type. This can be done iteratively with the following code:
 
-%TODO: code
+\begin{lstlisting}
+var query = {Track: {$exists: true}};
+
+db.train.find(query).forEach(function(doc) {
+
+  // Cast Track to be Int32 type
+  doc.Track = new NumberInt(doc.Track);
+
+  // Write changes to MongoDB
+  db.train.save(doc);
+
+});
+\end{lstlisting}
 
@@
 for handling enums is by using either the String or 32-bit Integer types. As the String type allows more expressive
-values and can be indexed to have integer-like performance, it would be most appropriate to change the field to be
+values, and can be indexed to have integer-like performance, it would be most appropriate to change the field to be
 represented as a String type, mapping the existing numbers to appropriate strings. As the existing \textbf{LoadEmpty}
@@
 
-%TODO: code
+\begin{lstlisting}
+// Change LoadEmpty = 1 to 'loaded'
+db.train.update(
+
+  // query
+  { "LoadEmpty" : 1 },
+
+  // update
+  { $set: {LoadEmpty: 'loaded'} },
+
+  // options
+  {
+    "multi" : true,   // Update all documents
+    "upsert" : false  // Insert a new document, if no existing document match the query
+  }
+);
+
+// Change LoadEmpty = 0 to 'empty'
+db.train.update(
+
+  // query
+  { "LoadEmpty" : 0 },
+
+  // update
+  { $set: {LoadEmpty: 'empty'} },
+
+  // options
+  {
+    "multi" : true,   // Update all documents
+    "upsert" : false  // Insert a new document, if no existing document match the query
+  }
+);
+
+// Change LoadEmpty = -1 to 'unknown'
+db.train.update(
+
+  // query
+  { "LoadEmpty" : -1 },
+
+  // update
+  { $set: {LoadEmpty: 'unknown'} },
+
+  // options
+  {
+    "multi" : true,   // Update all documents
+    "upsert" : false  // Insert a new document, if no existing document match the query
+  }
+);
+\end{lstlisting}
 
 Note that three different updates need to be performed on the collection, for each different value of \textbf{LoadEmpty},
-as filtered out by the query argument. As each update will need to query the collection to retreive a subset of documents
+as filtered out by the query argument. As each update will need to query the collection to retrieve a subset of documents
 to operate on, it would be wise to create a MongoDB index on the \textbf{LoadEmpty} key. This single-key index can be
@@
 
-%TODO: code
+\begin{lstlisting}
+// Create an ascending index on LoadEmpty field
+db.train.createIndex({LoadEmpty: 1});
+\end{lstlisting}
 
@@
 
-%TODO: JSON example
+\begin{lstlisting}
+{
+  "_id" : ObjectId("5561e94530f853aef8bbfd0c"),
+  "Time" : 1094468487.0999999046325684,
+  "Speed" : 4.4783997535999998,
+  "LoadEmpty" : -1.0000000000000000,
+  "km" : 5.0096998214999999,
+  "Lat" : -20.6665750019999983,
+  "Lon" : 117.1280816499999986,
+  "Track" : 2.0000000000000000,
+  "SND1" : 0.0118295024810000,
+  "SND2" : -0.0173957574360000,
+  "SND3" : -0.0295975576070000,
+  "SND4" : -0.0327308348410000,
+  "CouplerForce" : 23.1114991600000010,
+  "LateralAccel" : 0.0033679292537000,
+  "AccLeft" : 0.0000000000000000,
+  "AccRight" : 0.0000000000000000,
+  "BounceFront" : -0.0104506661800000,
+  "BounceRear" : -0.0234966575210000,
+  "RockFront" : 0.0445603373210000,
+  "RockRear" : 0.0122018001710000
+}
+\end{lstlisting}
 
@@
 
-%TODO: grouped JSON
+\begin{lstlisting}
+{
+  "_id" : ObjectId("555d30ca33d9d7607772920f"),
+  "Time" : 1094468487.0999999046325684,
+  "Speed" : 4.4783997535999998,
+  "LoadEmpty" : "unknown",
+  "km" : 5.0096998214999999,
+  "Track" : 2,
+  "CouplerForce" : 23.1114991600000010,
+  "LateralAccel" : 0.0033679292537000,
+  "SND" : {
+    "SND1" : 0.0118295024810000,
+    "SND2" : -0.0173957574360000,
+    "SND3" : -0.0295975576070000,
+    "SND4" : -0.0327308348410000
+  },
+  "Loc" : {
+    "Lat" : -20.6665750019999983,
+    "Lon" : 117.1280816499999986
+  },
+  "Bounce" : {
+    "Front" : -0.0104506661800000,
+    "Rear" : -0.0234966575210000
+  },
+  "Rock" : {
+    "Front" : 0.0445603373210000,
+    "Rear" : 0.0122018001710000
+  },
+  "Acc" : {
+    "Left" : 0.0000000000000000,
+    "Right" : 0.0000000000000000
+  }
+}
+\end{lstlisting}
 
@@
 
-%TODO: resturcturing code
+\begin{lstlisting}
+var updateDocsFunc = function(doc) {
+  // Add location nested doc
+  doc.Loc = {
+    Lat: doc.Lat,
+    Lon: doc.Lon
+  };
+
+  // Add Bounce nested doc
+  doc.Bounce = {
+    Front: doc.BounceFront,
+    Rear: doc.BounceRear
+  };
+
+  // Add Rock nested doc
+  doc.Rock = {
+    Front: doc.RockFront,
+    Rear: doc.RockRear
+  };
+
+  // Add SND nested doc
+  doc.SND = {
+    SND1: doc.SND1,
+    SND2: doc.SND2,
+    SND3: doc.SND3,
+    SND4: doc.SND4
+  };
+
+  // Add Acc nested doc
+  doc.Acc = {
+    Left: doc.AccLeft,
+    Right: doc.AccRight
+  };
+
+  // Delete old values
+  delete doc.Lat;
+  delete doc.Lon;
+  delete doc.BounceFront;
+  delete doc.BounceRear;
+  delete doc.RockFront;
+  delete doc.RockRear;
+  delete doc.SND1;
+  delete doc.SND2;
+  delete doc.SND3;
+  delete doc.SND4;
+  delete doc.AccLeft;
+  delete doc.AccRight;
+
+  // Write changes
+  db.train.save(doc);
+};
+
+// Iteratively perform operations on all records
+db.train.find().snapshot().forEach(updateDocsFunc);
+\end{lstlisting}
 
@@
 
-%TODO: code for Loc
-
-Finally, the collection contains documents in a similar structure to:
-
-%TODO: final JSON structure
+\begin{lstlisting}
+// Create a geospatial index on Loc nested document
+db.train.createIndex({Loc: '2d'});
+\end{lstlisting}
 
@@
 
-\section{Evaluation and Discussion} % (fold)
-\label{sec:evaluation_and_discussion}
+\section{Discussion} % (fold)
+\label{sec:discussion}
 

### 2015-05-25 baddata.js
Decomposed queries.js into sub script files

@@ -1,0 +1,2 @@
+// Find all documents stating a value less than 0 for KM travelled (these are obviously wrong)
+db.train.find({km: {$lt: 0}});

### 2015-05-25 fixtypes.js

@@ -1,0 +1,69 @@
+/*
+Fixes the default types set by mongoimport.
+
+- Explicitly casts all 'Track' values to Int32 type.
+- Updates all 'LoadEmpty' values to their String equivalents.
+*/
+
+
+// Iteratively casts all 'Track' values to Int32 type
+db.train.find({Track: {$exists: true}}).forEach(function(doc) {
+  // Cast Track to be a Int32 type
+  doc.Track = new NumberInt(doc.Track);
+
+  // Write changes to DB
+  db.train.save(doc);
+});
+
+
+// Update all 'LoadEmpty' values to String equivalents
+
+
+// Create an ascending index on LoadEmpty field for faster queries
+db.train.createIndex({LoadEmpty: 1});
+
+// Change LoadEmpty = 1 to 'loaded'
+db.train.update(
+  // query
+  { "LoadEmpty" : 1 },
+
+  // update
+  { $set: {LoadEmpty: 'loaded'} },
+
+  // options
+  {
+    "multi" : true,   // update all documents
+    "upsert" : false  // insert a new document, if no existing document match the query
+  }
+);
+
+// Change LoadEmpty = 0 to 'empty'
+db.train.update(
+  // query
+  { "LoadEmpty" : 0 },
+
+  // update
+  { $set: {LoadEmpty: 'empty'} },
+
+  // options
+  {
+    "multi" : true,   // update all documents
+    "upsert" : false  // insert a new document, if no existing document match the query
+  }
+);
+
+// Change LoadEmpty = -1 to 'unknown'
+db.train.update(
+  // query
+  { "LoadEmpty" : -1 },
+
+  // update
+  { $set: {LoadEmpty: 'unknown'} },
+
+  // options
+  {
+    "multi" : true,   // update all documents
+    "upsert" : false  // insert a new document, if no existing document match the query
+  }
+);
+

### 2015-05-25 mapreduce.js

@@ -1,0 +1,33 @@
+/*
+Runs a basic map reduce to get the total distance travelled on each track.
+Note that it excludes where km value is given as less than 0.
+ */
+
+// Select the Track and km fields from all documents
+var mapFunc = function() {
+  emit(this.Track, this.km);
+};
+
+// Reduces the km values from all mapped documents
+var reduceFunc = function(keyTrack, kms) {
+  return Array.sum(kms);
+};
+
+// Grabs the average of the reduced km values
+var finalizeFunc = function(key, reducedVal) {
+  reducedVal.avg = reducedVal.qty / reducedVal.count;
+
+  return reducedVal;
+};
+
+// Execute the map reduce with the specified functions
+db.train.mapReduce(
+  mapFunc,
+  reduceFunc,
+  {
+    out: 'total_distance',      // Stores output into new 'total_distance' collection
+    query: { km: { $gt: 0 }},   // Excludes bad (< 0) km values
+    finalize: finalizeFunc
+  }
+);
+

### 2015-05-25 queries.js

@@ -1,155 +1,0 @@
-// Command to insert CSV with headers to Mongo:
-//mongoimport -d railway -c train --type csv --file ~/Downloads/test.csv --headerline --ignoreBlanks
-
-/*
-Runs a basic map reduce to get the total distance travelled on each track
-Note that it excludes where km value is given as less than 0.
- */
-function totalDistanceMR() {
-  var mapFunc = function() {
-    emit(this.Track, this.km);
-  };
-
-  var reduceFunc = function(keyTrack, kms) {
-    return Array.sum(kms);
-  };
-
-  var finalizeFunc = function(key, reducedVal) {
-    reducedVal.avg = reducedVal.qty / reducedVal.count;
-
-    return reducedVal;
-  };
-
-  db.train.mapReduce(
-    mapFunc,
-    reduceFunc,
-    {
-      out: 'total_distance',
-      query: { km: { $gt: 0 }},
-      finalize: finalizeFunc
-    }
-  );
-}
-
-function fixTypes() {
-  db.train.find({Track: {$exists: true}}).forEach(function(doc) {
-    // Cast Track to be a Int32 type
-    doc.Track = new NumberInt(doc.Track);
-
-    // Write changes to DB
-    db.train.save(doc);
-  });
-
-  // Change LoadEmpty = 1 to 'loaded'
-  db.train.update(
-    // query
-    { "LoadEmpty" : 1 },
-
-    // update
-    { $set: {LoadEmpty: 'loaded'} },
-
-    // options
-    {
-      "multi" : true,  // update only one document
-      "upsert" : false  // insert a new document, if no existing document match the query
-    }
-  );
-
-  // Change LoadEmpty = 0 to 'empty'
-  db.train.update(
-    // query
-    { "LoadEmpty" : 0 },
-
-    // update
-    { $set: {LoadEmpty: 'empty'} },
-
-    // options
-    {
-      "multi" : true,  // update only one document
-      "upsert" : false  // insert a new document, if no existing document match the query
-    }
-  );
-
-  // Change LoadEmpty = -1 to 'unknown'
-  db.train.update(
-    // query
-    { "LoadEmpty" : -1 },
-
-    // update
-    { $set: {LoadEmpty: 'unknown'} },
-
-    // options
-    {
-      "multi" : true,  // update only one document
-      "upsert" : false  // insert a new document, if no existing document match the query
[…]

### 2015-05-25 restructure.js

@@ -1,0 +1,63 @@
+/*
+Iteratively restructure default documents by grouping related fields into subdocuments.
+ */
+var updateDocsFunc = function(doc) {
+  // Add location nested doc
+  doc.Loc = {
+    Lat: doc.Lat,
+    Lon: doc.Lon
+  };
+
+  // Add Bounce nested doc
+  doc.Bounce = {
+    Front: doc.BounceFront,
+    Rear: doc.BounceRear
+  };
+
+  // Add Rock nested doc
+  doc.Rock = {
+    Front: doc.RockFront,
+    Rear: doc.RockRear
+  };
+
+  // Add SND nested doc
+  doc.SND = {
+    SND1: doc.SND1,
+    SND2: doc.SND2,
+    SND3: doc.SND3,
+    SND4: doc.SND4
+  };
+
+  // Add Acc nested doc
+  doc.Acc = {
+    Left: doc.AccLeft,
+    Right: doc.AccRight
+  };
+
+  // Delete old values
+  delete doc.Lat;
+  delete doc.Lon;
+  delete doc.BounceFront;
+  delete doc.BounceRear;
+  delete doc.RockFront;
+  delete doc.RockRear;
+  delete doc.SND1;
+  delete doc.SND2;
+  delete doc.SND3;
+  delete doc.SND4;
+  delete doc.AccLeft;
+  delete doc.AccRight;
+
+  // Write changes
+  db.train.save(doc);
+};
+
+// Iteratively perform restructuring
+db.train.find().forEach(updateDocsFunc);
+
+// Create a geospatial index on Loc nested document
+db.train.createIndex({Loc: '2d'});
+
+// Create an ascending index on km field
+db.train.createIndex({km: 1});
+

### 2015-05-25 insertCSVtoMongo.js
Removed old CSV to Mongo script

@@ -1,145 +1,0 @@
-/*
- * * * * * * * * * * * * * * * * *
- * * * SCRIPT INITIALISATION * * *
- * * * * * * * * * * * * * * * * *
- */
-
-// Dependencies
-CSVtoJSONConverter = require('csvtojson').core.Converter;
-fs = require('fs');
-MongoClient = require('mongodb').MongoClient;
-
-
-// CLI arg constants
-var NUM_ARGS = 3,
-    CSV_ARG  = 2;
-
-
-// Connection constants
-var HOST    = '127.0.0.1',
-    PORT    = '27017',
-    DB_NAME = 'railway',
-    MONGO_URL = 'mongodb://' + HOST + ':' + PORT + '/' + DB_NAME;
-
-
-// Intermediate JSON path
-var JSON_OUTPUT_PATH = '/Volumes/RAM Disk/csv_to_json_mongo.json';
-
-
-
-
-/*
- * * * * * * * * * * * * * * * * * * *
- * * * START OF HELPER FUNCTIONS * * *
- * * * * * * * * * * * * * * * * * * *
- */
-
-/*
-Given CLI args, checks them for what is expected
- */
-var checkArgs = function(args){
-  // Check for expected number of args
-  if (args.length != NUM_ARGS) {
-    console.error('usage: node insertCSVToMongo.js [path/to/file.csv]');
-    process.exit(1);
-  }
-
-  // Check for read access on the CSV path
-  fs.access(args[CSV_ARG], fs.R_OK, function(error) {
-    if (error) {
-      console.error('Cannot access "' + args[CSV_ARG] + '" for reading: ' + error);
-      process.exit(1);
-    }
-  });
-
-  // Check for existing file on the JSON path
-  fs.access(JSON_OUTPUT_PATH, fs.F_OK, function(error) {
-    if (!error) {
-      console.error(JSON_OUTPUT_PATH + ' already exists. Please choose another path.');
-      process.exit(1);
-    }
-  });
-};
-
-
-/*
-Given a path to a valid CSV file and a callback, converts CSV to JSON and inserts to DB using callback.
-
-Callback is run after conversion is finished.
- */
-var JSONFromCSV = function(csvPath, callback) {
-  var convertParams = {
-    'constructResult': false,   // Don't place JSON in memory, as input data can be large
-    'ignoreEmpty':     true,    // Ignore empty columns, as we don't need them in Mongo
[…]

### 2015-05-25 baddata.js
Added updates for removing of bad data in baddata.js

@@ -1,2 +1,32 @@
+/*
+Examples of MongoDB updates to remove bad data from database.
+ */
+
 // Find all documents stating a value less than 0 for KM travelled (these are obviously wrong)
-db.train.find({km: {$lt: 0}});
+// Unset the km value, which essentially removes the key-value pair from the document, while retaining all other pairs.
+db.train.update(
+  { km: {$lt: 0} },
+  { $unset: {km: 1} },
+  { multi: true }
+);
+
+// Unset all bad speed values
+db.train.update(
+  { speed: {$lt: 0} },
+  { $unset: {speed: 1} },
+  { multi: true }
+);
+
+// Unset all bad LoadEmpty values
+db.train.update(
+  { LoadEmpty: {$not: {$in: ['unknown', 'empty', 'loaded']} } },
+  { $unset: {LoadEmpty: 1} },
+  { multi: true }
+);
+
+// Unset all bad Track values
+db.train.update(
+  { $or: [ {Track: {$lt: 0}}, {Track: {$gt: 213841}} ] },
+  { $unset: {Track: 1} },
+  { multi: true }
+);

### 2015-05-25 restructure.js

@@ -1,6 +1,7 @@
 /*
 Iteratively restructure default documents by grouping related fields into subdocuments.
  */
+
 var updateDocsFunc = function(doc) {
   // Add location nested doc
   doc.Loc = {
@@ -60,4 +61,3 @@ db.train.createIndex({Loc: '2d'});
 
 // Create an ascending index on km field
 db.train.createIndex({km: 1});
-

### 2015-05-25 paper/paper.tex
Finished off content of implementation section

@@
 \lstdefinelanguage{JavaScript}{
-  keywords={typeof, new, true, false, catch, function, delete, return, null, catch, switch, var, if, in, while, do, else, case, break},
+  keywords={typeof, new, true, false, catch, function, delete, return, undefined, null, catch, switch, var, if, in, while, do, else, case, break},
   keywordstyle=\color{blue}\bfseries,
@@
   showspaces=false,
-   numbers=left,
-   numberstyle=\footnotesize,
-   numbersep=9pt,
+  numbers=none,
   tabsize=2,
@@
 import the data into. The \texttt{--headerline} flag asks Mongo to use the CSV header to automatically generate BSON
-keys for each document. So each row in the CSV file will be converted to a MongoDB document, and stored in the specified
+keys for each document. The \texttt{--ignoreBlanks} flag asks Mongo not to set empty CSV values to the empty string. Instead,
+these key-value pairs are left out of their respective documents.
+So each row in the CSV file will be converted to a MongoDB document, and stored in the specified
 database collection. By default, MongoDB tries to guess the data type for each value, with numbers represented as Doubles.
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[caption=Performing an iterative operation on database collection ``train''.]
 var query = {Track: {$exists: true}};
@@
 db.train.find(query).forEach(function(doc) {
-
   // Cast Track to be Int32 type
@@
   db.train.save(doc);
-
 });
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[caption=Performing asynchronous \texttt{update()} operations on ``train'' collection.]
 // Change LoadEmpty = 1 to 'loaded'
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[caption=Creating an index on ``LoadEmpty'' key in ``train'' collection.]
 // Create an ascending index on LoadEmpty field
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[caption=Original ``flat'' document structure.]
 {
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[caption=Improved multi-layered document structure.]
 {
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[caption=Restructuring of documents in ``train'' collection using iterative operation.]
 var updateDocsFunc = function(doc) {
@@
 simply get ignored from the operation. For example, if a collection contains 100 documents, where only four of those
-documents contain the \textbf{isAdmin} key and an operation is called that depends on the key \textbf{isAdmin}, only
+documents contain the \textbf{isAdmin} key, and an operation is called that depends on the key \textbf{isAdmin}, only
 those four documents within the collection will be operated on. This allows completely differently structured documents
-to be contained within the same collection (however, in a well-structured dataset, multiple collections would be used).
+to be contained within the same collection (however, in a well-designed Mongo database, multiple collections would be used).
 
@@
 
+Often, the sensor readings received by the IRT team contains noisy data, defined by impossible values. For example, often
+the kilometre tracking sensor can return distance values consisting of negative numbers. These are deemed to be bad
+readings, and thus should not be included in any queries that are performed on the data. Hence, this data needs to be
+excluded from the overall dataset in the MongoDB collection.
+
+To exclude such data, it is possible to use MongoDB's provided \texttt{update()} function, which allows the selection
+of noisy data based on a specified predicate. After data has been selected, \texttt{update()} allows the use of the
+\texttt{\$unset} operator to remove specified fields from documents. This allows the other sensor readings in the same
+document to persist in the database, while only removing those readings specified as noisy by the predicate.
+
+This is performed through the following code:
+
+\begin{lstlisting}[caption=MongoDB \texttt{update()} functions to remove data deemed to be noisy from existing documents.]
+// Unset the bad km values (those with values less than 0)
+db.train.update(
+  { km: {$lt: 0} },       // predicate
+  { $unset: {km: 1} },    // operation to perform
+  { multi: true }         // option to specify operation on _all_ matching documents
+);
+
+// Unset all bad speed values (those with values less than 0)
+db.train.update(
+  { speed: {$lt: 0} },
+  { $unset: {speed: 1} },
+  { multi: true }
+);
+
+// Unset all bad LoadEmpty values (those not in valid enum values)
+db.train.update(
+  { LoadEmpty: {$not: {$in: ['unknown', 'empty', 'loaded']} } },
+  { $unset: {LoadEmpty: 1} },
+  { multi: true }
+);
+
+// Unset all bad Track values (those not in range of [0,213841])
+db.train.update(
+  { $or: [ {Track: {$lt: 0}}, {Track: {$gt: 213841}} ] },
+  { $unset: {Track: 1} },
+  { multi: true }
+);
+\end{lstlisting}
+
 % subsection filtering_bad_data (end)
+
+\subsection{Ad-hoc adding of additional sensors} % (fold)
+\label{sub:adding_additional_sensors}
+
+In the case of the IRT project, it is very possible that additional sensors could be added at any future stage. For a
+relational database, this would require modification to the table schemas, and possible restructuring operations
+to be performed on existing tables to set values in the columns for the additional sensors. In MongoDB, as documents are
+schema-less, new datasets containing new sensor's data can be serialised as BSON and inserted into existing collections
+at will, without requiring any modifications to the underlying database.
+
+As stated previously, differently structured documents can co-exist in the same collection. Hence, adding datasets
+with additional sensors will simply result in documents being inserted that contain a different number of fields than
+those documents that were previously stored in that collection.
+
+For testing, an additional dataset has been created, based off the original dataset, with the addition of two extra
+sensor fields: \textbf{Temp} and \textbf{Weight}. Documents with the addition of these new sensors simply have two
+additional key-value pairs that are not present in those documents consisting of data from the original dataset. Any
+queries performed that depend on either the \textbf{Temp} or \textbf{Weight} keys will simply only use those documents
+where such values exist for these keys, thus ignoring documents from the original dataset, as touched on in~\sectref{sub:handling_missing_values}.
+
+An example document containing these new fields is shown as follows:
+
+\begin{lstlisting}[caption=Example of a document with extra sensor fields that do not exist in original dataset.]
+{
+  "_id" : ObjectId("556291145946aa793572b39a"),
+  "Time" : 1.09447e+09,
+  "Speed" : 4.4639997480000000,
+  "LoadEmpty" : "unknown",
+  "km" : 5.0096998220000000,
+  "Track" : 2,
+  "CouplerForce" : 23.1613006199999987,
+  "Weight" : 2380,   // new pair
+  "Temp" : 77,       // new pair
+  "Loc" : {
+    "Lat" : -20.6665759599999994,
+    "Lon" : 117.1280811000000028
+  },
+  "Bounce" : {
+    "Front" : -0.0303000000000000,
+    "Rear" : -0.0029600000000000
+  },
+  "Rock" : {
+    "Front" : 0.0383000000000000,
+    "Rear" : 0.0062500000000000
+  },
+  "SND" : {
+    "SND1" : -0.0111000000000000,
+    "SND2" : 0.0001660000000000,
+    "SND3" : undefined,
+    "SND4" : -0.0494000000000000
+  },
+  "Acc" : {
+    "Left" : 0.0000000000000000,
+    "Right" : 0.0000000000000000
+  }
+}
+\end{lstlisting}
+
+% subsection adding_additional_sensors (end)
 
@@
 \label{sec:discussion}
+
+As most of what has been shown in~\sectref{sec:implementation}, properly setting up the MongoDB database for use with the IRT
+project is not a hard task and only requires relatively trivial MongoDB operations to be performed.
 

### 2015-05-25 paper/paper.tex
Added Discussion and Conclusion section content

@@
 structure on a particular collection's key, allowing all documents within that collection with that key to be accessed
-in $O(log n)$ time, as opposed to $O(n)$ time if unindexed. MongoDB also supports numerous types of indexes, such as
+in $O(log(n))$ time, as opposed to $O(n)$ time if unindexed. MongoDB also supports numerous types of indexes, such as
 single-key, multi-key, geospatial, and hashed indexes.
@@
 As most of what has been shown in~\sectref{sec:implementation}, properly setting up the MongoDB database for use with the IRT
-project is not a hard task and only requires relatively trivial MongoDB operations to be performed.
+project is not a hard task and only requires relatively trivial MongoDB operations to be performed. The overall design
+decisions for the project is the more difficult part of the implementation, and is highly dependant on the use-case
+for the database. Hence, more time generally would be spent on the design of the MongoDB database, and use of the
+mongo shell, or appropriate MongoDB driver/API for a specific programming language, can be used to structure everything.
+
+As noted in~\sectref{sec:mongo_overview}, MongoDB allows for much more advanced operations to be performed, however
+most of the available functions are easy to learn and offer varying amounts of power over given database collections,
+depending how many options the programmer makes use of.
+
+Coming from a object-oriented and functional software engineering background, I found interacting with MongoDB through its provided JavaScript shell
+a very easy task, and far more intuitive than using declarative SQL on a traditional relational database. However, this is merely
+anecdotal, and for a database user with much experience with SQL and the relational data model, using MongoDB, and
+similarly functioning NoSQL databases, may prove to be a difficult task; particularly for those without a lot of imperative programming
+experience. However, as a counter-argument, for basic interaction with MongoDB, scripts can easily be written by
+following the API documentations without requiring a thorough understanding of the underlying programming languages.
+
+A further difference that may catch many users of relational database, is the lack of joins, as noted in~\sectref{ssub:joins}.
+As MongoDB does not natively support multi-document joins, and actively encourages alternative strategies, users will
+need to factor this into their database design decisions, using either nested documents or document references. While this
+reportedly improves query performance, it may add further complexity to the database design for relational database
+users.
+
+Apart from the implementation of the overall design of the IRT MongoDB database, MongoDB also provides functions for
+advanced queries and calculations to be performed on database data. One example of this would be the use of the map
+reduce programming model for calculations on MongoDB data. Shown below is an example of MongoDB's map reduce for
+calculating the total and average distances travelled by trains on each different railway track. This requires more
+advanced used of host programming language features.
+
+\begin{lstlisting}[caption=A basic map reduce implementation to get total and average distance travelled by trains on each track.]
+// Select the Track and km fields from all documents
+var mapFunc = function() {
+  emit(this.Track, this.km);
+};
+
+// Reduces the km values from all mapped documents
+var reduceFunc = function(keyTrack, kms) {
+  return Array.sum(kms);
+};
+
+// Grabs the average of the reduced km values
+var finalizeFunc = function(key, reducedVal) {
+  reducedVal.avg = reducedVal.qty / reducedVal.count;
+
+  return reducedVal;
+};
+
+// Execute the map reduce with the specified functions
+db.train.mapReduce(
+  mapFunc,
+  reduceFunc,
+  {
+    out: 'total_distance',      // Stores output into new 'total_distance' collection
+    query: { km: { $gt: 0 }},   // Excludes bad (< 0) km values
+    finalize: finalizeFunc
+  }
+);
+\end{lstlisting}
 
@@
 \label{sec:conclusion}
+
+In conclusion, the Monash University Institute of Railway Technology's railway data project has been shown to be possibly
+managed on a non-relational database, MongoDB. We have highlighted the differences in overall database design that have
+to be taken into consideration when implementing the dataset on a non-relational database, and ways the database
+handles special cases that have been proving to be a problem on their current relational model.
+
+Overall, the implementation of their dataset on a non-relational database has been positive, without any major issues that
+would impact the database design in a negative way. The one aspect that would need to be looked at further, would be a
+quantitative evaluation of given queries that are commonly performed on their current relational database, compared with
+the equivalent queries in MongoDB. We have shown that it is very possible and beneficial to use MongoDB with the IRT
+project data, however performance is an important aspect and should be looked into further to allow an informed decision
+to be made.
 

### 2015-05-25 paper/paper.tex
Minor proof-reading changes and font change

@@
 \usepackage{poltakmacros}           % Personal macro package (https://gist.github.com/poltak/6ed59e8f30c9c399143f)
+\usepackage[T1]{fontenc}
+\usepackage[scaled=.8]{sourcecodepro}
 \usepackage{listings}
@@
 
-While relational database management systems (RDBMS) have been somewhat of a ``go-to'' solution for for a number of years
+While relational database management systems (RDBMS) have been somewhat of a ``go-to'' solution for a number of years
 for general data storage and management problems that many application developers face, we have noted a recent rise
 in the use of non-relational data management tools~\cite{padhy2011rdbms}. Most of these tools have traditionally fallen into the domain of
-big data analytics, with platforms such as the Hadoop ecosystem~\footnote{https://hadoop.apache.org/} being notably popular.
+big data analytics, with platforms such as the Hadoop ecosystem\footnote{https://hadoop.apache.org/} being notably popular.
 However, outside of the domain of big data, looking more at general purpose data storage and management, what are now
@@
 Codd in 1970~\cite{codd1970relational}, unlike the more commonly used RDBMS technologies, such as
-MySQL~\footnote{https://www.mysql.com/}. Being free of the strictness the relational model enforces on its data allows
+MySQL\footnote{https://www.mysql.com/}. Being free of the strictness that the relational model enforces on its data allows
 NoSQL databases to focus less on the overall structure of data, and more on factors such as scalability and
@@
 
-In this paper, we will look at the use of MongoDB~\footnote{https://www.mongodb.org/}, a popular NoSQL database solution,
+In this paper, we will look at the use of MongoDB\footnote{https://www.mongodb.org/}, a popular NoSQL database,
 as a solution for a case study based upon a railway data problem using data from Monash University's Institute of Railway
 Technology (IRT). An overview of the case study in question will be given in~\sectref{sec:case_study_overview}. A small
-overview of the MongoDB database will be given in~\sectref{sec:mongo_overview}. Implementation and
-discussion details will be given in~\sectref{sec:implementation} and~\sectref{sec:discussion}, respectively,
+overview of the MongoDB database will be given in~\sectref{sec:mongo_overview}. Implementation details and
+discussion will be given in~\sectref{sec:implementation} and~\sectref{sec:discussion}, respectively,
 before concluding in~\sectref{sec:conclusion}.
@@
 The case study that is being looked at involves a project that has been worked at at the Monash University Institute of
-Railway technology (IRT). The project involves trains the operate in the Pilbara region of Western Australia, taking
+Railway Technology (IRT). The project involves trains the operate in the Pilbara region of Western Australia, taking
 ore and minerals from loading points at mines to a specified unloading port. The data that is being dealt with comes from
-numerous categories of sensors being placed on certain specialised train cars to record data monitoring track and train
-car conditions~\cite{darby2003development,darby2005track}.
+numerous categories of sensors being placed on certain specialised train cars to record the data for monitoring the conditions
+for track and train cars~\cite{darby2003development,darby2005track}.
 
-Data currently gets unloaded at sent back in large batches to remote servers once the car completes a trip and pulls back
-into port~\cite{thomas2012taking}. The project currently makes use of a RDBMS solution to manage data storage, however
+Data currently gets unloaded and sent back in large batches to remote servers once the car completes a trip and pulls back
+into port~\cite{thomas2012taking}. The project currently makes use of an RDBMS solution to manage data storage, however
 this solution is faced with many problems that currently require painful work-arounds. The most obvious of which involves
@@
 be guaranteed to be returned from such a sensor. Hence, data received by the remote server is often inconsistently
-structured, and thus the data has to go through a series of preprocessing work arounds to fit in within the strict
-schema that the RDBMS expects. As the only guaranteed consistency in the data that gets received is the timestamp and
-geocoordinates, the strictly structured relational model is not an appropriate solution.
+structured, and thus the data has to go through a series of preprocessing work-arounds to fit in within the strict
+schema that the RDBMS expects. As the only guaranteed consistency in the received data is the timestamp and
+geocoordinates, the strictly structured relational model is currently deemed to not be an appropriate solution.
 
@@
 replace the current system with an appropriately tested solution. While the scope of that project is much larger than
-what will be covered in this paper, we will look at the differences in what is possible when attempting to use a NoSQL
+what will be covered in this report, we will look at the differences in what is possible when attempting to use a NoSQL
 database for the data storage and management.
@@
 its data~\cite{parker2013comparing}. Data within Mongo is stored using JSON-like structured documents, called
-BSON~\cite{bsons6:online}, which allows for strongly typed data fields and binary data. BSON documents are then stored
+BSON~\cite{bsons6:online}, which allow for strongly typed data fields and binary data. BSON documents are then stored
 within ``collections'' which reside in a given ``database''~\cite{parker2013comparing}. To relate back to the relational
@@
 \centering
-\caption{MongoDB conceptual structures as they relate to the relational model concepts.}
+\caption{MongoDB concepts as they relate to RDBMS concepts.}
 \label{tab:mongo_to_relational}
@@
 coexist freely. BSON documents generally follow some sort of predefined structure for that collection, to maintain
-data structure sanity, for the user's sake, however none of this is enforced.
+data structure sanity for the user's sake, however none of this is enforced.
 
-This sort of freedom in terms of structure allows for similarly structured documents in database collections, where
-differences may only be found in that certain documents may miss keys that are present in other documents. This is
+In practise, this freedom in terms of structure allows for similarly structured documents in database collections, where
+differences may only be found in that certain documents may be missing keys that are present in other documents, or inversely,
+documents may exist with extra keys not found elsewhere. This is
 highly rational in the way that, depending on the use-case, data may not be present for certain fields all the time.
@@
 
-MongoDB allows a lot of the similar type functionality in queries and data manipulation that is also expected in
-relational databases. Generally interactions with data are performed through JavaScript functions on collections, or cursors
+MongoDB allows a lot of the similar type functionality, in terms of queries and data manipulation, that is also expected in
+relational databases. Generally, interactions with data are performed through JavaScript functions on collections, or cursors
 (subsets of documents within a collection), allowing general querying operations through use of the \texttt{find()} function,
@@
 languages, including JavaScript, Python, and Java. As well as providing drivers, MongoDB also provides a default shell,
-called the \textit{mongo shell}, which allows general JavaScript code to be written to interact with the standard MongoDB
+called the \textit{mongo shell}, which allows general JavaScript code to be written and interpreted to interact with the standard MongoDB
 API and perform ad-hoc queries. No official graphical user interface exists to interact with MongoDB, however there
@@
 functions are user defined operations to perform on documents in collections to return results. These include general
-aggregate operation models, such as MapReduce, and often involve grouping documents by given keys and various queries
+aggregate operation models, such as map reduce, and often involve grouping documents by given keys and various queries
 to limit the amount of documents to process.
@@
 Instead, MongoDB allows two main ways of relating different documents to each other. These are via referencing and
-nested documents. Nested documents are the main recommended way of join multiple documents, unless the documents have
-many-to-many relationships. To allow nested documents, a BSON value can store a whole other document, or arrays of documents.
-Example use-cases of where nested documents are appropriate include documents that have natural ``child'' documents that
-only it will ever need to reference. For example, a blog post document may have many nested comment documents as the
-comments are specific to a given blog post. For documents where many things may refer to it, it is recommended to use
+nested documents. Nested documents are the recommended way of ``joining'' multiple documents, unless the documents have
+many-to-many relationships. To allow nested documents, a BSON value can store a whole document, or arrays of documents.
+Example use-cases of where nested documents are appropriate include documents that have natural child documents that
+only the parent will ever need to reference. For example, a blog post document may have many nested comment documents as the
+comments are specific to a given blog post. For documents where many other documents may refer to it, it is recommended to use
 references via BSON values containing other document IDs.
+
+For performance purposes, nested documents can be accessed in a single query, while referenced documents require multiple
+queries depending on the amount of document references needed to be traversed.
 
@@
 in $O(log(n))$ time, as opposed to $O(n)$ time if unindexed. MongoDB also supports numerous types of indexes, such as
-single-key, multi-key, geospatial, and hashed indexes.
+single-key, multi-key, geospatial, and hashed indexes. Examples of creating indexes will be given in~\sectref{sec:implementation}.
 
@@
 redundancy and availability. Replication in Mongo is managed using a primary instance of data, with secondaries replicating
-the data in the primary using operation logs (\textit{oplogs}) to mimic any operations on data that were performed in the
-primary. Replicas have fault tolerance in the way that if the primary fails, an election is started to promote any of the
+the data in the primary using operation logs (\textit{oplogs}) to mimic any operations on performed by the primary.
+Replicas have fault tolerance in the way that if the primary fails, an election is started to promote any of the
 secondaries to become a primary. In general, a client will only interact with a primary, however operations can also be
-specified to be performed on specific secondaries. However, as replication data is asynchronously managed, any reads
-performed specifically on secondaries may not give up-to-date data.
+specified to be performed on specific secondaries. However, as replication data between the primary and secondaries is asynchronously managed, any reads
+performed specifically on secondaries may not return up-to-date data.
 
@@
 
-Sharding in MongoDB allows for horizontal scaling of a Mongo database, where a given database or collection is shared
-among different ``shards'', or multiple MongoDB instances, generally running on separate machines. Sharding is performed
+Sharding in MongoDB allows for the horizontal scaling of a Mongo database. Sharding is where a given database, or collection, is shared
+among different ``shards'', or MongoDB instances, generally running on separate machines. Sharding is performed
 on an indexed document key, dividing documents containing that key based on either range or computed hashes. Ranged-based
-sharding is generally recommended where data is generally expected to be accessed in some sort of sequence, while hash-based
-sharding offers a guaranteed random distribution of data across shards, and more balanced equally-sized shards.
+sharding is recommended where data is generally expected to be accessed in some sort of sequence, while hash-based
+sharding offers a guaranteed random distribution of data across shards, thus giving more balanced, equally-sized shards.
 
@@
 Implementation of the database for the given Monash IRT data was all performed using MongoDB version 3.0.3, the latest
-version as of 2015-05-18. Interaction with the MongoDB database was all performed using the mongo shell and corresponding
-official NodeJS MongoDB driver.
+version as of 2015-05-18. Interaction with the MongoDB database was all performed using JavaScript code via the mongo shell.
 
@@
 The test data used for this case study was acquired from the Monash IRT team, and was an extract of real data they have
-received from the railway sensors. As only one set of data was given, simulated datasets based upon the extract have
+received from the railway sensors. As only one set of data was provided, simulated datasets based upon the extract have
 also been created to simulate the type of inconsistencies the IRT team are facing, including differently structured
@@
 \begin{table}[h]
-\caption{The different types of data received from the Monash University Institute of Railway Technology team.}
+\caption{The different types of sensor data received from the Monash University Institute of Railway Technology team.}
 \label{tab:irt_data}
@@
 
-This dataset contains data from all the possible sensors. As previously stated, the datasets received are not always
+This dataset contains data from all the currently installed sensors. As previously stated, the datasets received are not always
 consistent, hence not all of these fields will always be present, or may be present and contain incorrect data readings.
@@
 format can easily handle all of the given data types. All given types can be handled using the BSON primitive types of
-Double and 32-bit Integer for floating point and integer data, respectively.
+Double and 32-bit Integer for floating point and integer values, respectively.
 
@@
 
-The \texttt{-d} and \texttt{-c} flags enable the specifying of database and collection within that database in which to
+The \texttt{-d} and \texttt{-c} flags enable the specifying of database and collection in which to
 import the data into. The \texttt{--headerline} flag asks Mongo to use the CSV header to automatically generate BSON
@@
 these key-value pairs are left out of their respective documents.
-So each row in the CSV file will be converted to a MongoDB document, and stored in the specified
-database collection. By default, MongoDB tries to guess the data type for each value, with numbers represented as Doubles.
+By default, MongoDB tries to guess the data type for each value, with numbers represented as Doubles.
 Of course, storing every document as a flat container for different floating point values is not optimal, and pre-processing
@@
 
-As the documents now contain appropriately typed key-value pairs, further modifications to the documents can be performed
+As the documents now contain appropriately typed values, further modifications to the documents can be performed
 to better structure the documents. By default, the CSV importer structures the documents generated from CSV rows as flat
@@
 
-Further restructuring can be done to group the \textbf{Bounce}, \textbf{Rock}, and \textbf{Acc} values into sub-documents
+Further restructuring can be done to group the \textbf{Bounce}, \textbf{Rock}, \textbf{Lat}, \textbf{Lon}, and \textbf{Acc} values into sub-documents
 containing key-value pairs for each, finally ending up with the following document structure:
@@
 
-These restructurings could be performed with the following code, in mongo shell for example:
+These restructurings could be performed iteratively using the following code, in mongo shell for example:
 
-\begin{lstlisting}[caption=Restructuring of documents in ``train'' collection using iterative operation.]
+\begin{lstlisting}[caption=Restructuring of documents in ``train'' collection using an iteratively applied operation.]
 var updateDocsFunc = function(doc) {
@@
 
-Finally, one more restructure could be performed on the existing document structure, which is grouping the \textbf{Lat}
-and \textbf{Lon} values into a single sub-document, which could be named \textbf{Loc}. This further enables future
-geospatial indexes being more easily created on the location data, for location-dependent queries and operations.
+Finally, as we have now grouped the \textbf{Lat}
+and \textbf{Lon} values into the single sub-document, \textbf{Loc}, this enables us to create a geospatial index on this
+sub-document.
+This geospatial index will allow for faster location-dependent queries and operations.
 This restructuring can be performed by the following code:
@@
 
-For the test datasets missing data for particular sensors, MongoDB handles these cases. The datasets include
+For the test datasets missing data for particular sensors, MongoDB natively handles these cases. The possible datasets include
 CSV files with different value layouts, and often missing values in random rows, apart from the time and location fields
-as they are guaranteed.
+which are guaranteed.
 
@@
 not have any ordering of its keys, documents are still generated with the appropriate data. For datasets with missing
-values in certain rows, the BSON document that is generated simply does not contain a key-value pair for the missed
-attribute. This is opposed to setting the key to a null value, or default value. Of course, a user may generate keys
+values in certain rows, the BSON document that is generated simply does not contain a key-value pair for the missing
+attribute. This is opposed to setting the key to a null or default value. Of course, a user is free to generate keys
 for the missing values and point them to null or default values, however there is no advantage to the extra processing.
@@
 Overall, the implementation of their dataset on a non-relational database has been positive, without any major issues that
-would impact the database design in a negative way. The one aspect that would need to be looked at further, would be a
+would impact the database design in any negative way. The one aspect that would need to be looked at further, would be a
 quantitative evaluation of given queries that are commonly performed on their current relational database, compared with
